# Predicting the Number of Bike Shares
In this analysis we would like to predict the count of bike rentals in Washington DC

## 1. Getting Data
### Data source
We downloaded the dataset "bikesharing.csv" from https://code.datasciencedojo.com/datasciencedojo/datasets/tree/master/Bike%20Sharing.

### Loading the data
In a first step the dataset is imported to R and stored in the data.frame *d.bike*:
```{r}
d.bike <- read.csv("bikesharing.csv", header=TRUE)
```

### Describing the dataset
```{r}
str(d.bike)
```

The dataset contains hourly information about a day and its weather conditions in 17379 observations of 17 variables. In the following, the individual attributes will be explained:

* *instant*: record index
* *dteday*: Date
* *season*: 1 = spring, 2 = summer, 3 = fall, 4 = winter
* *yr*: Year; 0: 2011, 1: 2012
* *mnth*: Month; 1 to 12
* *hr*: Hour; 0 to 23
* *holiday*: 0 = no holiday, 1 = holiday
* *weekday*: Day of the week; 1 to 7
* *workingday*: 0 = weekend or holiday, 1 = working day
* *weathersit*:
  + 1: Clear, Few clouds, Partly cloudy, Partly cloudy
  + 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
  + 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
  + 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog 
* *temp*: Normalized temperature in Celsius; The values are derived via (t-t_min)/(t_max-t_min),   t_min=-8, t_max=+39;
* *atemp*:  Normalized feeling temperature in Celsius; Normalized   feeling temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-16, t_max=+50
* *hum*: Normalized relative humidity; The values are divided to 100 (max); 
* *windspeed*: Normalized windspeed ; The values are divided to 67 (max)
* *casual*: number of non-registered  (eg. tourists)
* *registered*: number of registered users (eg. workers that use bikes daily)
* *cnt*: number of total rentals

### Feature Selection
We decide to delete the feature *instant* and *dteday*. *instant* is a record index that has been artificially created and *dteday* is the date of the record. Therefore both will have no relevant effect on *cnt*.

```{r}
d.bike <- subset(d.bike, select=-c(instant, dteday))
```

### Checking the data
```{r}
head(d.bike)
tail(d.bike)
```

As it looks like the data set was imported completely. To check if there are any missing values (not available, NA) in this dataset we count the number of NAs in the data set.
```{r}
sum(is.na(d.bike))
mean(is.na(d.bike))
```

The record contains no rows with missing data. Now we are ready for the further analysis steps.

### Creating training and test data set


## 2. Explorative Data Analysis
Because ggplot allows elegant data visualization with the same syntax for different types of data models, we decide to use ggplot exclusively for graphical representations as far as possible. Therefore we import ggplot2.
```{r}
library(ggplot2)
```

### Count
In the very first step we would like to look at the distribution of the response variable count. For this purpose we create a histogram.
```{r}
ggplot(data = d.bike, aes(x=cnt)) +
  geom_histogram(bins=30, colour="black", ylab="Frequency") + xlab("Count") + ylab("Frequency")
mean(d.bike$cnt)
sd(d.bike$cnt)
```
As we can see, this is a right skewed distribution. So in most hours - the number of bike rentals are hourly observations - between 0 and 30 bikes (the most left bar) are rented.

### Season
If we want to visualize the effect of the *season* as a categorical variable on the number of bike rentals it is a good idea to use a box plot.
```{r}
ggplot(data = d.bike, aes(group=season, y = cnt, x = as.factor(season))) +
  geom_boxplot() + xlab("Season") + ylab("Count")
```

As the box plot shows, season 1 (spring) is particularly different from season 2, 3 and 4 (summer, autumn and winter). The median is lowest in spring, even lower than in winter. In summer, autumn and winter, the medians differ only marginally.  Since the medians differ partially, one can assume a slight effect. However, we want to examine this more closely.

With an Analysis of Variance (ANOVA) we check whether at least one level of Seasons has a significant effect on the number of bike rentals. The first step is to create a model that does not consider any effect from *season*. With an F-test we compare then both models with each other.

```{r}
lm.season.1 <- lm(cnt ~ as.factor(season), data = d.bike)
lm.season.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.season.0, lm.season.1)
```

As we can see from the p-value there is strong evidence that at least one level in *season* has a strong effect on the number of bike rentals. The lower value of the residual sums of squares in lm.season.1 also indicates that the more complex model has smaller residuals than the less complex model lm.season.0.

### Year
Since we also consider year to be a categorical variable, a box plot is again a good choice. For all categorical variables we will use a boxplot for the graphical representation in the following.

```{r}
ggplot(data = d.bike, aes(group=yr, y = cnt, x = yr)) +
  geom_boxplot()
```

At first glance we can see that the number of bike rentals has increased in year 1 (2012). The median is above the median of year 0 (2011).

```{r}
lm.year.1 <- lm(cnt ~ as.factor(yr), data = d.bike)
lm.year.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.year.0, lm.year.1)
```


### Month
```{r}
ggplot(data = d.bike, aes(group=mnth, y = cnt, x = as.factor(mnth))) +
  geom_boxplot() + scale_x_discrete(limits=seq(1,12)) + xlab("Month") + ylab("Count")
```

Here we see that the bike rental increased during the summer months May to September or when the temperatures were more pleasant.

```{r}
lm.month.1 <- lm(cnt ~ as.factor(mnth), data = d.bike)
lm.month.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.month.0, lm.month.1)
```

### Hour
```{r}
ggplot(data = d.bike, aes(group=hr, y = cnt, x = as.factor(hr))) +
  geom_boxplot() + scale_x_discrete(limits=seq(1,23)) + xlab("Hour") + ylab("Count")
```

The same applies to the time of day and hour. Bikes were most often rented in the morning between 7 and 8 pm. and in the evening between 5 and 6 am. The same flow also applies to the median.

```{r}
lm.hr.1 <- lm(cnt ~ as.factor(hr), data = d.bike)
lm.hr.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.hr.0, lm.hr.1)
```

### Holiday
```{r}
ggplot(data = d.bike, aes(group=holiday, y = cnt, x = as.factor(holiday))) +
  geom_boxplot() + xlab("Holiday") + ylab("Count")
```

During the holiday period, fewer bikes are rented according to the boxplot diagram. However, the difference is very pleasant compared to the normal working period. The holiday period median is close to the work time median.

```{r}
lm.holiday.1 <- lm(cnt ~ as.factor(holiday), data = d.bike)
lm.holiday.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.holiday.0, lm.holiday.1)
```

### Weekday
```{r}
ggplot(data = d.bike, aes(group=weekday, y= cnt, x = as.factor(weekday))) +
  geom_boxplot() + xlab("Weekday") + ylab("Count")
```

During all seven days of the week, the bike rental is almost constant. On Monday and Sunday there are less bikes rented. The median is lowest on Monday and highest on Saturday. On all other days the median is similar.  

```{r}
lm.weekday.1 <- lm(cnt ~ as.factor(weekday), data = d.bike)
lm.weekday.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.weekday.0, lm.weekday.1)
```

### Working day
```{r}
ggplot(data = d.bike, aes(group=workingday, y = cnt, x = as.factor(workingday))) +
  geom_boxplot() + xlab("Working day") + ylab("Count")
```

On working or not-working days, the count of the bike rentals are almost equal. This is also identical with the both medians.

```{r}
lm.workingday.1 <- lm(cnt ~ as.factor(workingday), data = d.bike)
lm.workingday.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.workingday.0, lm.workingday.1)
```

### Weather Situation
```{r}
ggplot(data = d.bike, aes(group=weathersit, y = cnt, x = as.factor(weathersit))) +
  geom_boxplot()  + xlab("Weather Situation") + ylab("Count")
```

If we look at the weather divided on a scale between 1 and 4 as shown in the boxplot, we see that the bike rental is steadily decreasing from 1 to 4. The value 1 is the good weather and 4 the bad weather. Whereby from the value 3 away, probably already bad weather, the renting goes down faster. An identical flow can also be observed for the medians, starting from the first box plot 1 to 4.

```{r}
lm.weathersit.1 <- lm(cnt ~ as.factor(weathersit), data = d.bike)
lm.weathersit.0 <- lm(cnt ~ 1, data = d.bike)
anova(lm.weathersit.0, lm.weathersit.1)
```

### Temperature

```{r}
lm.temp.1 <- lm(cnt ~ temp, data = d.bike)
summary(lm.temp.1)
``` 


```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = temp)) +
  geom_point()  + xlab("Normalized Temp") + ylab("Count") + geom_abline(intercept=lm.temp.1$coefficients[1], slope=lm.temp.1$coefficients[2], colour="red", size=1)
``` 


### Feeling Temperature

```{r}
lm.atemp.1 <- lm(cnt ~ atemp, data = d.bike)
summary(lm.atemp.1)
``` 

```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = atemp)) +
  geom_point() + xlab("Normalized Feeling Temp") + ylab("Count") + geom_abline(intercept=lm.atemp.1$coefficients[1], slope=lm.atemp.1$coefficients[2], colour="red", size=1)
```

### Humidity

```{r}
lm.hum.1 <- lm(cnt ~ hum, data = d.bike)
summary(lm.hum.1)
``` 

```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = hum)) +
  geom_point() + xlab("Humidity") + ylab("Count") + geom_abline(intercept=lm.hum.1$coefficients[1], slope=lm.hum.1$coefficients[2], colour="red", size=1)
```

### Windspeed

```{r}
lm.windspeed.1 <- lm(cnt ~ windspeed, data = d.bike)
summary(lm.windspeed.1)
``` 

```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = windspeed)) +
  geom_point() + xlab("Windspeed") + ylab("Count") + geom_abline(intercept=lm.windspeed.1$coefficients[1], slope=lm.windspeed.1$coefficients[2], colour="red", size=1)
```

### Casual Users

```{r}
lm.casual.1 <- lm(cnt ~ casual, data = d.bike)
summary(lm.casual.1)
``` 

```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = casual)) +
  geom_point() + xlab("Casual Users") + ylab("Count") + geom_abline(intercept=lm.casual.1$coefficients[1], slope=lm.casual.1$coefficients[2], colour="red", size=1)
```

### Registered Users

```{r}
lm.registered.1 <- lm(cnt ~ registered, data = d.bike)
summary(lm.registered.1)
``` 

```{r}
ggplot(data = d.bike, mapping = aes(y = cnt, x = registered)) +
  geom_point() + xlab("Registered Users") + ylab("Count")+ geom_abline(intercept=lm.registered.1$coefficients[1], slope=lm.registered.1$coefficients[2], colour="red", size=1)
```

### Hypothesis
In this section, the insights gained from the exploratory data analysis and the consideration of the individual attributes should help to form a hypothesis about what the final model might look like. This hypothesis is the starting point for developing the model.



## 3. Regression Analysis

### Model developing 

### Linear Regression

### Non-linear Regression

### Poisson Regression

### Final model

```{r}
poi.bike.1 <- glm(cnt ~ as.factor(season) + as.factor(yr) + as.factor(mnth) + as.factor(hr) + as.factor(holiday) + as.factor(weekday) + as.factor(workingday) + as.factor(weathersit) + temp + atemp + hum + windspeed, family = "poisson", data = d.bike)
summary(poi.bike.1)
```



### Logistic Regression

### Cross Validation (simple versus complex Model)


## 4. Trees

## 5. Support Vector Machines

## 6. Neural Networks

## 7. Conclusion (Comparing all models in one table)