---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Predicting the Quality of Wine
## 1. Getting Data
### Data source
We downloaded the dataset "winequalityN.csv" from https://www.kaggle.com/rajyellow46/wine-quality.

### Loading the data
In a first step the dataset is imported to R and stored in the data.frame *d.wine*:
```{r}
d.wine <- read.csv("winequalityN.csv", header=TRUE)
```

### Describing the dataset
```{r}
str(d.wine)
```

The dataset contains content information of different red and white wines in 6497 observations of 13 columns. In the following, the individual attributes will be explained:

* **type**: categorial predictor with 2 levels white/red that describes whether the wine is a red or white wine.
* **fixed.acidity**: continous predictor that describes the amount of acids that are solid and do not evaporate easily.
* **volatile.acidity**: continous predictor that describes the amount of acids that can lead to a vinegar like taste.
* **citric.acid**: continous predictor that describes the amount of acids that can add freshness and flavor to wines.
* **residual.sugar**: continous predictor that describes the amount of sugar remaining after fermentation. Wines with greater than 45 grams/liter are considered sweet.
* **chlorides**: continous predictor that describes the amount of salt in the wine.
* **free.sulfur**.dioxide: continous predictor that describes the amount of the free form of sulphur dioxide (SO2). It prevents microbial growth and the oxidation of wine.
* **total.sulfur.dioxide**: continous predictor that describes the amount of the free and the bound form of sulphur dioxide (S02). A concentration greater than 50 ppm becomes evident in nose and mouth.
* **density**: continous predictor that describes the density of the water in the wine.
* **pH**: continous predictor that describes how acidic or basic a wine is on a scale of 0 (very acidic) and 14 (very basic). Most wines have a pH value between 3 and 4.
* **sulphates**: continous predictor that describes the amount of the wine additive which can contribute to sulfur dioxide gas (S02) levels.
* **alcohol**: continous predictor that describes the percent alcohol content of the wine.
* **quality**: categorical response variable with 10 levels between 0 and 10 that describes the wine quality.

### Checking the data
```{r}
head(d.wine)
tail(d.wine)
```
As it looks like the data set was imported completely. In row No 6494 there is an missing value (not avaiable, NA) in the **sulphates** column. Probably this is not the only one. Therefore we count the number of NAs in the data set.
```{r}
sum(is.na(d.wine))
mean(is.na(d.wine))
```
The complete data set contains 38 NA. These make up about 0.04% of the data set. We decide to delete the incomplete rows.
```{r}
d.wine <- na.omit(d.wine)
sum(is.na(d.wine))
```
The data set now contains only complete observations. Now we are ready for the further analysis steps.

## 2. Graphical Analysis
