# Predicting the Number of Bike Shares
In this analysis we would like to predict the count of bike rentals in Washington DC.
## 1. Getting Data
### Data source
We downloaded the dataset "bikesharing.csv" from https://code.datasciencedojo.com/datasciencedojo/datasets/tree/master/Bike%20Sharing.

### Loading the data
In a first step the dataset is imported to R and stored in the data.frame *d.bike*:
```{r}
d.bike <- read.csv("bikesharing.csv", header=TRUE)
```

### Describing the dataset
```{r}
str(d.bike)
```

The dataset contains hourly information about a day and its weather conditions in 17379 observations of 17 variables. In the following, the individual attributes will be explained:

* *instant*: record index
* *dteday*: Date
* *season*: 1 = spring, 2 = summer, 3 = fall, 4 = winter
* *yr*: Year; 0: 2011, 1: 2012
* *mnth*: Month; 1 to 12
* *hr*: Hour; 0 to 23
* *holiday*: 0 = no holiday, 1 = holiday
* *weekday*: Day of the week; 1 to 7
* *workingday*: 0 = weekend or holiday, 1 = working day
* *weathersit*:
  + 1: Clear, Few clouds, Partly cloudy, Partly cloudy
  + 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
  + 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
  + 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog 
* *temp*: Normalized temperature in Celsius; The values are derived via (t-t_min)/(t_max-t_min),   t_min=-8, t_max=+39;
* *atemp*:  Normalized feeling temperature in Celsius; Normalized   feeling temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-16, t_max=+50
* *hum*: Normalized relative humidity; The values are divided to 100 (max); 
* *windspeed*: Normalized windspeed ; The values are divided to 67 (max)
* *casual*: number of non-registered  (eg. tourists)
* *registered*: number of registered users (eg. workers that use bikes daily)
* *cnt*: number of total rentals

### Checking the data
```{r}
head(d.bike)
tail(d.bike)
```
As it looks like the data set was imported completely. To check if there are any missing values (not available, NA) in this dataset we count the number of NAs in the data set.
```{r}
sum(is.na(d.bike))
mean(is.na(d.bike))
```
The record contains no rows with missing data. Now we are ready for the further analysis steps.

## 2. Graphical Analysis
